%%% response-reviewer1.tex ---
%%
%% Filename: response-reviewer1.tex
%% Description:
%% Author: George Tsoulas
%% Maintainer:
%% Created: Fri Dec  2 10:58:03 2016 (+0000)
%% Version:
%% Package-Requires: ()
%% Last-Updated:
%%           By:
%%     Update #: 0
%% URL:
%% Doc URL:
%% Keywords:
%% Compatibility:
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%% Commentary:
%%
%%
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%% Change Log:
%%
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% This program is free software; you can redistribute it and/or
%% modify it under the terms of the GNU General Public License as
%% published by the Free Software Foundation; either version 3, or
%% (at your option) any later version.
%%
%% This program is distributed in the hope that it will be useful,
%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
%% General Public License for more details.
%%
%% You should have received a copy of the GNU General Public License
%% along with this program; see the file COPYING.  If not, write to
%% the Free Software Foundation, Inc., 51 Franklin Street, Fifth
%% Floor, Boston, MA 02110-1301, USA.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%% Code:

\documentclass[11pt]{article}
\usepackage{linguex}
\usepackage{todonotes}
\usepackage{natbib}
\bibliographystyle{chicago}

\newcommand{\citeposs}[1]{\citeauthor{#1}'s (\citeyear{#1})}

\begin{document}
\title{Response to third reviewer's comments}
\author{this is the review starting \textit{this is a well written paper\ldots}}
\date{December 2016}

\maketitle


The reviewer's insightful and helpful comments have allowed us to clarify a number of issues in the paper and have led to several modifications both major and minor in the presentation and argument. We thus would like to begin by thanking the reviewer for the time and effort that (s)he put in commenting on the paper. We are pleased that we have been able to incorporate the vast majority of the comments.

More specifically, We have followed the reviewer's suggestions and incorporated earlier discussions of the idea of differential spell-out, although we have limited ourselves to the areas where this was most relevant to the concerns of the paper, e.g. the issues of scattered deletion in \cite{boskovic-nunes:2007}. However, it is important to note that while the operation is descriptively the same, namely the spelling out different bits of material in head and tail copies, the derivation of the rich typology wh-constructions is substantially different from say, scattered deletion in Macedonian and Slavic clitic clusters -- there is not a clear crosslinguistic PF-constraint that cuts through wh-constructions. Of course, it would be extremely interesting if we could somehow predict the surface configuration of a wh-question in a particular language in terms of PF-constraints in that language. We have to admit, however, that articulating these constraints across the wide range of languages we consider is beyond the scope of our current ability.

More importantly, we would like to address here two issues of substance that the reviewer raises and offer some clarification:

First, the reviewer raises an issue concerning the connection that we make between the analysis of wh-scope and the \citeposs{Beghelli-Stowell97} feature-based system for scope.  The reviewer suggests that it may be problematic, in the framework of the paper, that choice function indefinites (which we take to be wide-scope indefinites) and Q-particles do not behave alike.  The reviewer recognises that this is not in itself a problem but goes on to point out that  Q-particles do not behave exactly like  existential quantifiers. This is seen as problematic in the context of our suggestion (modelled on \citeposs{kratzer:05} work) that there may be an existential position in the set of scope-related positions in the clausal spine, corresponding to the position where $\exists$-closure applies. Strictly speaking, this is a correct observation as in the semantic literature it is claimed that choice-function indefinites can also take lower scope. However, the comparison with Q-particles in this respect is not really warranted because Q-particles are not taken to be existential quantifiers.  Although we have not taken an explicit stance on this question (but we have also not challenged it, so implicitly we are going with it), the analysis of Q-particles due to \citet{hagstrom:1998} and later, especially \citet{cable:2007} takes them to be variables over choice functions which can be bound either by existential closure or an interrogative C resulting in an indefinite reading of question reading, respectively. What this really means is that the QP approach proposes that all QP clusters are wh-indefinites and result in question meaning only when bound by interrogative C (the propositional set-former). Given that semantic analyses of closure specifically take it to be an unrestricted operation that can take place at different points \citep[see most importantly,][and subsequent work]{Reinhart:1997,winter:1997a}  It is to be expected that, if this is so, $\exists$Ps may be found at different points along the clausal spine. This is in fact to be expected on a variety of grounds, theoretically, the suggestion that phasal nodes are accompanied by a functional shell with significant similarities across phase heads becomes increasingly commonplace.  Empirically, there have been suggestions that elements like a DistP are indeed found within DPs \citep[e.g.][]{etxepare:14,tsoulas-woods:15}. It then follows that, if anything, this behaviour supports our proposal since it underlines a parallel between choice function indefinites which can be ``bound'' at wide or narrow scope levels by closure operations (this is because the closure operation is not restricted) and SSO as applied to QR.

The reviewer also asks why, under this account, QR would be clause bound unlike wh-movement. This is, of course, a very good question and as we have pointed out, there is really no good explicit account in the literature that assumes the classic version of the rule for this behaviour of QR. Conversely, the lack of clause-boundedness of wh-movement can be straightforwardly explained by the cyclicity of wh-movement. In so, then one would expect a formulation for QR along the lines that successive cyclicity is \textbf{not avaialble} in QR. We have not incorporated an explicit discussion of this point in the paper for reasons of space and to keep things more coherent, due to the speculative nature of the issue. Having said this, we believe there is potentially a good answer to this question that derives directly from the system. Our approach here is that the locality of QR issue is solved immediately under \citeposs{Beghelli-Stowell:97} system and our reinterpretation of its mechanisms in the following way: given that the effects of QR are subsumed under AGREE (and subsequent movement if needed) between a functional head and a DP it follows that QR will be restricted by the same locality constraints as AGREE. Therefore, a lack of crossclausal AGREE relations would preclude successive cyclicty, resulting in the clauseboundedness effects we see. A problem would only arise in a situation where there are no scopal functional structure in a subordinate clause and there is in the matrix.  We know of no such cases and the reason why they would be ruled out anyway is the independent principle of \textit{Phase Impenetrability Condition}.  In other words, by the time the higher scope heads have been merged the lower DPs would have been subject to TRANSFER, fixing embedded scope.

\bibliography{scope-glossa,fbib}
\end{document}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% response-reviewer1.tex ends here
